Formalization of Typed Lua

### 2015-06-10 doc/papers/paper/paper.tex
finishing rewrite of section on projection types for dls paper

At this point, using $q$ or $r$ as an r-value will project
their types into either $\Integer \cup \Nil$ or $\Integer \cup \String$. But the more interesting case is inside the
clauses of the {\bf if} statement. The particular uses the rule {\sc T-IF-FNIL-PROJ} below, which uses flow typing:
\[
\begin{array}{c}
\mylabel{T-IF-FNIL-PROJ}\\
\dfrac{\begin{array}{c}
       \env_{1}(id) = \pi_{i}^{x} \;\;\; \penv(x) = S\\
       closeall(\env_{1}), \penv[x \mapsto fpt(S,\Nil,i)] \vdash s_{1}, \env_{2} \\
       closeall(\env_{1}), \penv[x \mapsto gpt(S,\Nil,i)] \vdash s_{2}, \env_{3} \\
       \env_{4} = closeset(\env_{1}, fav(s_{1}) \cup fav(s_{2})) \\
       \env_{5} = openset(\env_{4}, rv(s_{1}) \cup rv(s_{2}))
      \end{array}}
      {\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{5}}
\end{array}
\]

@@

In our example, $x$ would be bound to $\Integer \times \Integer \times \Nil{*}$ in the {\bf then} branch, and would be bound
to $\Nil \times \String \times \Nil{*}$ inside the {\bf else}
branch.

Thus, reading $q$ and $r$ projects $\pi_{1}^{x}$ to $\Integer$ and
$\pi_{2}^{x}$ to $\Integer$ inside the $\mathbf{if}$ branch,
but it projects $\pi_{1}^{x}$ to $\Nil$ and $\pi_{2}^{x}$ to $\String$
inside the $\mathbf{else}$ branch.
Outside the condition, $q$ and $r$ use the original projection, that is,
they project to $\Integer \cup \Nil$ and $\Integer \cup \String$, respectively.

>>>>

At this point, using $q$ or $r$ as an r-value will project
their types into either $\Integer \cup \Nil$ or $\Integer \cup \String$. But the more interesting case is inside the
branches of the {\bf if} statement. The particular uses the rule {\sc T-IF-FNIL-PROJ} below, which uses flow typing:
\[
\begin{array}{c}
\mylabel{T-IF-FNIL-PROJ}\\
\dfrac{\begin{array}{c}
       \env_{1}(id) = \pi_{i}^{x} \;\;\; \penv(x) = S\\
       closeall(\env_{1}), \penv[x \mapsto fpt(S,\Nil,i)] \vdash s_{1}, \env_{2} \\
       closeall(\env_{1}), \penv[x \mapsto gpt(S,\Nil,i)] \vdash s_{2}, \env_{3} \\
       \env_{4} = closeset(\env_{1}, fav(s_{1}) \cup fav(s_{2})) \\
       \env_{5} = openset(\env_{4}, rv(s_{1}) \cup rv(s_{2}))
      \end{array}}
      {\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{5}}
\end{array}
\]

@@

In our example, $x$ would be bound to $\Integer \times \Integer \times \Nil{*}$ in the {\bf then} branch, and would be bound
to $\Nil \times \String \times \Nil{*}$ inside the {\bf else}
branch. Now $q$ and $r$ project to $\Integer$ and $\Integer$
in the {\bf then} branch, and to $\Nil$ and $\String$ in the
{\bf else} branch.

A projection type in l-value position does not trigger
the projection, effectively forbidding assignment to a
variable that has a projection type. Allowing assignment
would be unsafe, as it could break the dependency among the
components of the union, as the following example shows:
\begin{verbatim}
  local q, r = idiv(a, b)
  r = "foo"
  if q then
    -- wrongly assume q and r are integers
  end
\end{verbatim}
